---
title: "Comparison of Chess Computer Development Throughout the Years"

execute: 
  echo: false
---

Chess has been around since the 6th century, and in the 20th century, the game once studied by humans was conquered by computers. This began with IBM's Deep Blue engine, which defeated World Champion Garry Kasparov in 1997. After this historic match, chess engines began to develop at an unprecedented rate, and in present times, they have outclassed humans in strategy, technique, and calculation.

To begin with, let's analyze how certain computers have scaled to the strength they possess today by comparing their growth with one of the greats, World Champion Magnus Carlsen.

```{r}
#| message: false
library(tidyverse)
library(bigchess)
```

```{r}
#| message: false
#| warning: false
computer1 <- read.pgn("pgn files/deep_fritz.pgn")
computer2 <- read.pgn("pgn files/dragon_by_komodo.pgn")
computer3 <- read.pgn("pgn files/rybka.pgn")
computer4 <- read.pgn("pgn files/stockfish.pgn")

# Sample data
data <- data.frame(
  Year = rep(2009:2023, 2),
  Rating = c(2805, 2810, 2835, 2837, 2868, 2876, 2893, 2882, 2842, 2843, 2862, 2882, 2864, 2870, 2859,
             2600, 2700, 2800, 2900, 3000, 3100, 3200, 3300, 3400, 3500, 3600, 3700, 3800, 3900, 4000),
  Entity = rep(c("Magnus Carlsen", "Stockfish"), each = 15)
)

# Create the plot
p <- ggplot(data, aes(x = Year, y = Rating, color = Entity)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  labs(title = "Rating Growth of Magnus Carlsen vs. Stockfish",
       x = "Year",
       y = "Rating",
       color = "Entity") +
  theme_minimal()

# Print the plot
print(p)


```

